---
title: "Homework 15"
author: "Aidan Fauth"
format: html
editor: visual
---

## Introduction

This file is the material from chapter 3 of *Getting Started with R*. I will be using data from the book that deals with apple production. I will be learning about various data manipulation methods in this chapter primarily using the dplyr and ggplot packages.

## Set up

```{r}
rm(list=ls())
library(tidyverse)
library(here)
```

Import the data set:

```{r}
compensation = read.csv(here("Data", "compensation.csv"))
glimpse(compensation) # take a look at the data
```

Get summary statistics for each of the compensation variables using the `summay()` function

```{r}
summary(compensation)
```

The book had Grazing as a factored variable, but mine is only character.

```{r}
compensation$Grazing = as.factor(compensation$Grazing)
levels(compensation$Grazing) #check them out
```

Now the summary output should be the same as the book's:

```{r}
summary(compensation)
```

## Using dplyr 'Verbs'

-   Important note: dpylr functions are supplied a data frame and then return a data frame.

The first argument to all dplyr functions is the data frame (in this case the compensation data.)

`select()` -- grabs cols. To see names of columns use `names(compensation)`.

```{r}
select(compensation, Fruit)
```

Selecting all columns except one:

```{r}
select(compensation, -Root) # add a minus sign
```

`slice()` -- grabs rows including one, a sequence, or a discontinuous set.

```{r}
slice(compensation,2) # one row
slice(compensation,2:10) # set of rows
slice(compensation,c(2,3,10)) # discontinuous set
```

`filter()` -- powerful subsetting

requires knowledge of logical and relational operators (boolean values.)

Practicing with \> first using the `with()` function which asks R to look at a data frame and perform a action, then stop looking.

```{r}
with(compensation, Fruit > 80) # check to see if the fruit value is greater than 80
```

Boolean values are what is passed to `filter()` and then it acts on them.

```{r}
# find the rows where Fruit > 80 and return it as a data frame
filter(compensation, Fruit > 80)
```

Selecting rows on multiple conditions:

```{r}
filter(compensation, Fruit > 80 | Fruit < 20)
```

Saving this information in a variable:

```{r}
lo_hi_fruit = filter(compensation,Fruit > 80 | Fruit < 20)
lo_hi_fruit # now print it
```

Commit this to memory: " assign the values returned by subsetting using filter to an object if you want to use them again" (66).

## Transforming

`mutate()` -- designates a new column name in the df with the tranformed data.

Will do this by overwritting the original working df.

```{r}
head(compensation) # what the df looks like now

# use mutate
compensation = mutate(compensation,logFruit = log(Fruit))

# what compensation looks like now
head(compensation)
```

## Sorting

`arrange()` -- sorts the rows

Increasing order of Fruit:

```{r}
arrange(compensation,Fruit)
```

## Mini-summary and two tips

Tip 1: composition of dplyr functions:

```{r}
select(filter(compensation, Fruit > 80), Root) # Root values from Fruit > 80
```

Tip 2: pipe -- '%\>%' read as "put the answer of the left-hand command into the function on the right"

Note: piping with dplyr starts with a df

```{r}
compensation %>% filter(Fruit > 80) %>% select(Root)
```

Read from left to right: 1) use the compensation data 2) filter based on Fruit \> 80 3) pass it to select and only return the Root column data.

## Summarization

General Outline:

1.  Declare data frame and grouping variable

2.  Math function to summarize the data

3.  Name for the returned values

4.  Make R use all the information

Method 1: Nested -- No Piping
```{r}
summarise(group_by(compensation, Grazing),meanFruit = mean(Fruit)) # you can save this as an object by assigning it a variable name (i.e mean_Fruit <-/mean_Fruit = )
```

Method 2: Pipe -- No Nesting
```{r}
compensation %>% group_by(Grazing) %>% summarize(meanFruit = mean(Fruit))
```


## Extending Summarization

including more summary statistics like `mean()`, `sd()`, and `median()`
```{r}
compensation %>% group_by(Grazing) %>% summarize(meanFruit = mean(Fruit), sdFruit = sd(Fruit), medFruit = median(Fruit)) # I added the median which was not included in the book example.
```

