---
title: "Biostats - Homework 20"
format: html
editor: visual
---

### Introduction

The objective of this assignment is to have you continue practicing the skills you have learned in Chapters 3 and 4 of Getting Started with R and reinforce the ideas we discussed in class regarding the Weissgerber et al. paper. Please complete all of the following steps:

1.  Write the code to set up your workspace.

```{r}
rm(list=ls())
library(tidyverse)
library(here)
```

2.  Read in the dataset called soil_data.csv and name it "soil."

```{r}
soil <-  read.csv(here("Data", "soil_data.csv"))
```

3.  Create a histogram for the SOM_percent column in the dataframe. Accept the default number of bins.

```{r}
ggplot(soil, aes(SOM_percent))+
  geom_histogram()
```

How would you describe the mean_SOM data based on viewing the histogram? Write your answer below:

The distribution is very right skewed so the mean is probably between 10 and 15 -- possibly around 13.

4.  Create a new dataframe called SOM by calculating for each forest the mean soil organic matter and the standard deviation of soil organic matter. Make sure to name variables you create "mean_SOM" and "sd_SOM" so that the next code chunk I give you works.

```{r}
SOM <- soil |>  
  group_by(Forest) |> 
  summarise(
    mean_SOM = mean(SOM_percent),
    sd_SOM = sd(SOM_percent))
```

5.  Now run the following code chunk which will create a bar plot of mean soil organic matter per forest, with standard deviation as error bars.

```{r}
ggplot(SOM, aes(Forest, mean_SOM))+
  geom_errorbar(aes(ymax = mean_SOM + sd_SOM, ymin = mean_SOM - sd_SOM), width = 0.2)+
  geom_col(fill = "white", color = "black")+
  ylab("Mean percent soil organic matter")+
  xlab("Forest name")
```

Based on what you see in the bar chart, how would you describe the pattern of SOM between the 4 different forests? Write your answer here:

It shows that the Kip Tract has the highest mean percent soil organic matter followed by Donnerville, South Hammond, and Degrasse. With these data, the larger means also have a larger error bar (these forests have a larger standard deviation/spread of SOM percentages.) It's hard to see if the error bars overlap because only the positive part of the error bar is shown, but it is likely a good amount of the error bars overlap, meaning that the means are within one standard deviation of each other (in numerical order.)

Now create a boxplot of the SOM_percent data. Use Forest as the x-axis variable and overlay the point data on top of the boxplot. Make the points blue and somewhat transparent. Call the x-axis "Forest name" and the y-axis "Percent soil organic matter" (hint: See homework 18...).

```{r}
ggplot(soil, aes(x = Forest, y = SOM_percent))+
  geom_boxplot()+
  geom_point(color = "blue", alpha = 0.5)+
  xlab("Forest name")+
  ylab("Percent soil organic matter")

```

Based on what you see in the box plot, write out your answers the following questions:

-   how would you describe the pattern of SOM between the 4 different forests? The kip tract has larger median SOM among the other forests and the other three have very similar medians, but Donnerville has a clear outlier. The other three forests have very narrow spread compared to the Kip Tract which had a large spread and is skewed right.

-   What do you notice that is different from how the data were displayed in the bar plot? The Donnerville center (this time the average) seems much larger in comparison and the spread is much larger as well. The bar graph makes the Donnerville data seem more similar to the Kip Tract distribution than the box plots do.

-   Which plot type is more informative? Explain your answer.

The box plots because they show the center, shape, and spread of the data and the median is also not effected by skew. The bar graph only shows the center and the spread, and the mean is effected by skew and especially the extreme outliers in the Donnerville data.

Look at na.rm = T

```{r}
soil |> summarise(mean_SOM = mean(SOM_percent), sd_SOM = sd(SOM_percent)) |> View()
```

```{r}
soil |> summarise(mean_N = mean(Nitrate_N_ppm, na.rm = T), sd_SOM = sd(Nitrate_N_ppm, na.rm = T)) |> View()
```
