---
title: "Homework 25"
author: "Aidan Fauth"
format: html
editor: visual
---

This homework covers the code from sections 5.6 to 5.7 in the GSWR book (pg 118-129).

## Set up

```{r}
rm(list = ls())
library(tidyverse)
library(here)
library(ggfortify)
```

## 5.6 "Analysis fo variance: the one-way ANOVA"

Final test: One way ANOVA -- uses a factor/categorical explanatory variable and a continuous response variable.

Question with two parts: 1) Is there a relationship between the predictor and response? 2) How do the groups compare to each other?

Load in the data set:

```{r}
daphnia <- read.csv(here("Data", "Daphniagrowth.csv"))
glimpse(daphnia)
```

Need to make parasite into a factor:

```{r}
daphnia$parasite <- as.factor(daphnia$parasite)
```

Plot the data:

```{r}
ggplot(daphnia, aes(x=parasite, y = growth.rate))+
  geom_boxplot()+
  theme_bw()+
  coord_flip() # if the names are mashed together
```

Interpretation:

There is clear differences between the groups.

The control has the highest median growth rate of about 1.2 mm/day.

There is an ordering to the other group: P. perplexa, M, bicuspidata, and P. ramosa from highest to lowest median growth rate


5.6.2 "Construct the ANOVA"

Use lm()

```{r}
model_grow <- lm(growth.rate ~ parasite, data = daphnia)
```


5.6.3 "Check Assumptions"

```{r}
autoplot(model_grow, smooth.colour = NA)
```


The normal quantile plot is potentially concerning, but we can still accept that the assumptions are met.


5.6.4 "Making an Inference from a One-Way ANOVA"

Use anova()

```{r}
anova(model_grow)
```

Output looks the same as regression because the are both linear models.

Null hypothesis: there is no difference in group means.

F-value: ratio of "between group variance" and "within-group" variance. Large F-value means a small p-value.


5.6.5 "Treatment Contrasts"

Look at summary table:

```{r}
summary(model_grow)
```

What is missing? treatment group called control -- its not missing its just labelled intercept.

R presents them in alphabetical order with the first group being the intercept.

Treatment contrasts illustrate differences between the "reference level" and the other groups.

In this example: the differences between the parasite groups and the control.

Means for the groups can also be found using `group_by()` and `summarise()`:

```{r}
sumDat <- daphnia |> 
  group_by(parasite) |> 
  summarise(meanGR = mean(growth.rate))
sumDat
```

Coefficients are the control + treatment contrast (in `summary()`) = group mean

When the control is the intercept, the p-values are useful, but be wary of "multiple testing problem"

Plot:

```{r}
ggplot(daphnia,aes(parasite, growth.rate, colour = parasite))+
  geom_point(size = 3, alpha = .5)+
  geom_point(data = sumDat, aes(parasite, meanGR), shape = "diamond", size = 7 )+
  theme_bw()+
  coord_flip()
```


