---
title: "Homework 21"
author: "Aidan Fauth"
format: html
editor: visual
---

## Introduction to statistics:

This homework is covering the beginng of chapter 5 of GSWR book.

## Set up

```{r}
rm(list=ls())
library(tidyverse)
library(here)
```

Import the data and take a look at the data:

```{r}
lady <- read.csv(here("Data", "ladybirds.csv"))
head(lady)
```

Convert the char data to factored data:

```{r}
lady$Habitat <- as.factor(lady$Habitat)
lady$Site <- as.factor(lady$Site)
lady$colour <- as.factor(lady$colour)
```

Check to see if everything looks right:

```{r}
head(lady)
```


## $\chi^2$ Contingency Table Analysis

Looking to see if there is an association between the two color morphs and two habitats.

First, organize the data for plotting and analysis:

```{r}
totals <- lady |> 
  group_by(Habitat, colour) |> 
  summarise(total.number = sum(number))
```

Make a figure of these data using a bar chart:

```{r}
ggplot(totals, aes(x = Habitat, y = total.number, fill = colour))+
  geom_bar(stat = "identity", position = "dodge")
```

This is the only situations a bar graph makes sense because there is no spread or shape to the data to examine -- it's just count data.

Some new things to add to ggplot:

fill = colour used for geometry like bars that have a fill color. Also, color = color would alter to the outline of the bars

geom_bar() syntax:

  stat = "identity" stops ggplot from calculating anything from the data it was given.
  
  position = "dodge" tells ggplot to put the habitat groups next to each other instead of having a stacked bar graph.
  

Change the colors on the bar graph to make more sense:

```{r}
ggplot(totals, aes(x = Habitat, y = total.number, fill = colour))+
  geom_bar(stat = "identity", position = "dodge")+
  scale_fill_manual(values = c(black = "black", red = "red")) # change the colors to represent the data
```

Interpreting the graphs

Just from the graphs there does seem to be an association between the habitat type and the more prevalent color morph.

We expect to reject the null hypothesis due to an uneven distribution of color morphs among the habitats.


Making the $\chi^2$ test

We need the data frame to be converted to a matrix.

```{r}
lady.mat <- xtabs(number ~ Habitat + colour, data = lady) 
# xtabs cross tabulates to convert the df to a matrix using a simple formula
```

Now the test:

```{r}
chisq.test(lady.mat)
```

Reject the null. We have discernible evidence ($\chi^2 = 19.1, df = 1, p = 1.239 * 10^{-5}$) that the number of black and red color morphs are unevenly distributed among the two habitats.


## Two-sample t-test

The two sample t-test is a comparison of two means. The predictor is categorical and the response is quantitative.

Assumptions:
1. Data are normally distributed
2. Equal variance

Fist graph the data: show both distributions of each category

Use two histograms:

First load in the ozone data:

```{r}
ozone <- read.csv(here("Data", "ozone 2024-10-23 23_43_15.csv"))
```

```{r}
head(ozone)
```

Convert char data to factors:

```{r}
ozone$Garden.location <- as.factor(ozone$Garden.location)
ozone$Garden.ID <- as.factor(ozone$Garden.ID)
```

```{r}
head(ozone)
```
Looks good.

Step 1: plot the data

Histograms:

```{r}
ggplot(ozone, aes(Ozone))+
  geom_histogram(binwidth = 10)+
  facet_wrap(~ Garden.location, ncol = 1)+ # ncol arranges them in column
  theme_bw()
```

There is a good amount of overlap, but the means could be different. We need the test to be sure.

Two-sample t-test:

```{r}
t.test(Ozone ~ Garden.location, data = ozone)
```

We reject the null due to the small p-value.

Note: Welch t-test is due to the fact that the equal variance assumption was not met (slightly different t-test was used instead)


